---
title: "Indirect Reference Interval Estimation using lab.mclust"
author: "Sandra Klawitter, Weihao Qiu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Indirect Reference Interval Estimation using lab.mclust}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The lab.mclust() function provides an indirect method for estimating reference intervals using Gaussian Finite Mixture Models via the mclust algorithm. This approach is designed to model multi-modal or non-normal laboratory data and isolate the non-pathological subpopulation.

## Usage
Below is a standard workflow for applying lab.mclust to indirect reference interval calculations.

```{r setup, message = FALSE}
library(reflimR.expand)
library(reflimR)
```

We can estimate reference limits for Alanine Aminotransferase (ALT) using the built-in dataset:

```{r example, eval = FALSE}
# Load example dataset from reflimR
data("livertests", package = "reflimR")

# Run lab.mclust on ALT values
alt_data <- livertests$ALT
res <- lab.mclust(alt_data)

# Print results
print(res)
```
## References

Hoffmann, G., Trulson, I., Adler, J., Klawitter, S., Holdenrieder, S., & Klawonn, F. (2026). Indirect methods for the verification of reference intervals in laboratory medicine. Critical reviews in clinical laboratory sciences, 1–17. Advance online publication. https://doi.org/10.1080/10408363.2026.2687513
