---
title: "Indirect Reference Interval Estimation using lab_mclust"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Indirect Reference Interval Estimation using lab_mclust}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction
The lab_mclust() function provides an indirect method for estimating reference intervals using Gaussian Finite Mixture Models via the mclust algorithm. Originally developed in the VeRIf repository, this approach is designed to model multi-modal or non-normal laboratory data and isolate the non-pathological subpopulation.

## Basic Usage
Below is a standard workflow for applying lab_mclust to indirect reference interval calculations.

```{r setup, message = FALSE}
library(reflimR.expand)
library(reflimR)
```
### Example with livertests Dataset
We can estimate reference limits for Alanine Aminotransferase (ALT) using the built-in dataset:

```{r example, eval = FALSE}
# Load example dataset from reflimR
data("livertests", package = "reflimR")

# Run lab_mclust on ALT values
alt_data <- livertests$ALT
res <- lab_mclust(alt_data)

# Print results
print(res)
```
## References

Algorithm original source: https://github.com/SandraKla/VeRIf/blob/main/functions.R