---
title: "reflimR.Expand: Advanced Reference Limit Estimation Using Routine Laboratory Data"

output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reflimR.Expand: Advanced Reference Limit Estimation Using Routine Laboratory Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4.5
)
```

1.  [Introduction](#introduction)
2.  [Installation](#installation)
3.  [Usage](#usage)
4.  [References](#references)

# Introduction <a name = "introduction"></a> {#introduction}

Reference intervals play a crucial role in the medical interpretation and statistical evaluation of laboratory results. By definition, reference intervals include the central 95% of results measured in non-diseased reference individuals [1, 2].

The *reflimR.expand* package provides a powerful extension to the existing *reflimR* package. While *reflimR* focuses on estimating reference intervals from laboratory data, *reflimR.expand* adds methods specifically designed to address challenges in real-world data, such as:

* handling values below the limit of detection (LOD), and
* applying sliding window techniques for age-dependent reference intervals.

# Installation <a name = "installation"></a> {#installation}

To use the *reflimR.expand* package, you must first install its dependency, *reflimR*, from CRAN. Open R and enter the following command in the console:

```{r, eval = FALSE}
install.packages("reflimR")
```

This will download and install the *reflimR* package from CRAN. Next, install the *reflimR.expand* package directly from GitHub using the *devtools* package:

```{r, eval = FALSE}
# If devtools is not installed yet:
install.packages("devtools")

# Install reflimR.expand from GitHub
devtools::install_github("SandraKla/reflimR.expand")
```

Once the installation is complete, load the package into your R session with:

```{r}
library(reflimR)
library(reflimR.expand)
```

# Usage <a name = "usage"></a> {#usage}

After installing and loading the package, you can use the main functions to compute reference intervals with enhanced methods:

## 1. Sliding Window Reference Limits

Biological parameters often change continuously across covariates such as age. Instead of creating arbitrary discrete age groups, `sliding.reflim()` applies a moving window to compute continuous local reference limits.

### Example: Continuous Age-Dependent Limits

```{r}
set.seed(123)

# Generate synthetic age data (18 to 80 years)
age <- runif(1000, min = 18, max = 80)

# Generate biomarker values with an age-dependent upward trend
biomarker <- 40 + 0.12 * age + rnorm(1000, mean = 0, sd = 2.5)

# Calculate sliding reference limits
res_sliding <- sliding.reflim(
  x = biomarker,
  covariate = age,
  window.size = 20,   # Span of covariate per window
  step.width = 2,     # Shift per iteration
  n.min = 39          # Minimum required observations per window
)

# Inspect output structure
head(res_sliding)
```

### Visualizing Results

The companion function `draw.sliding.reflims()` visualizes local reference limits. You can plot results aligned by local median covariate (`use.mean = FALSE`) or mean covariate (`use.mean = TRUE`):

```{r fig.width=6, fig.height=4.5}
# Aligning by median covariate value
draw.sliding.reflims(res_sliding, use.mean = FALSE)
```

```{r fig.width=6, fig.height=4.5}
# Aligning by mean covariate value
draw.sliding.reflims(res_sliding, use.mean = TRUE)
```

## 2. Weighted Sliding Window

For datasets with sparse regions or non-uniform age distributions, localized kernel weights can be applied via weighted sliding functions (`reflim_Weighted_Sliding.R`) to assign higher weights to observations near the center of the window.

## 3. Limit of Detection (LOD) Functions

Laboratory results close to analytical sensitivity thresholds often fall below the Limit of Detection (LOD). The LOD routines (`reflimLOD.r`) handle left-censored measurements without introducing bias caused by deleting low values or imputing zeros.

# References <a name = "references"></a> {#references}

1. Clinical and Laboratory Standards Institute document EP28-A3c—Defining, Establishing, and Verifying Reference Intervals in the Clinical Laboratory. 2010;ISBN 1-56238-682-4. <https://clsi.org/shop/standards/ep28/>.

2. Hoffmann, G., Klawitter, S., Trulson, I., Adler, J., Holdenrieder, S., & Klawonn, F. (2024). A Novel Tool for the Rapid and Transparent Verification of Reference Intervals in Clinical Laboratories. Journal of clinical medicine, 13(15), 4397. <https://doi.org/10.3390/jcm13154397>.
