---
title: "reflimR.Expand: Advanced Reference Limit Estimation Using Routine Laboratory Data"

output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reflimR.Expand: Advanced Reference Limit Estimation Using Routine Laboratory Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

1.  [Introduction](#introduction)
2.  [Installation](#installation)
3.  [Usage](#usage)
4.  [References](#references)

# Introduction <a name = "introduction"></a> {#introduction}

Reference intervals play a crucial role in the medical interpretation and statistical evaluation of laboratory results. By definition, reference intervals include the central 95% of results measured in non-diseased reference individuals [1, 2].

The *reflimR.expand* package provides a powerful extension to the existing *reflimR* package. While *reflimR* focuses on estimating reference intervals from laboratory data, *reflimR.expand* adds methods specifically designed to address challenges in real-world data, such as:

* handling values below the limit of detection (LOD), and
* applying sliding window techniques for age-dependent reference intervals.

# Installation <a name = "installation"></a> {#installation}

To use the *reflimR.expand* package, you must first install its dependency, *reflimR*, from CRAN. Open R and enter the following command in the console:

```{r, eval = FALSE}
install.packages("reflimR")
```

This will download and install the *reflimR* package from CRAN. Next, install the *reflimR.expand* package directly from GitHub using the *devtools* package:

```{r, eval = FALSE}
# If devtools is not installed yet:
install.packages("devtools")

# Install reflimR.expand from GitHub
devtools::install_github("SandraKla/reflimR.expand")
```

Once the installation is complete, load the package into your R session with:

```{r}
library(reflimR.expand)
```

# Usage <a name = "usage"></a> {#usage}

After installing and loading the package, you can use the main functions to compute reference intervals with enhanced methods:

* Use the Sliding Window functions to calculate reference intervals based on a sliding window approach, allowing for local estimation across covariates such as age.
* Use the reflimLOD functions to handle and adjust data values below the limit of detection (LOD).

# References <a name = "references"></a> {#references}

1. Clinical and Laboratory Standards Institute document EP28-A3c—Defining, Establishing, and Verifying Reference Intervals in the Clinical Laboratory. 2010;ISBN 1-56238-682-4. <https://clsi.org/shop/standards/ep28/>.

2. Hoffmann, G., Klawitter, S., Trulson, I., Adler, J., Holdenrieder, S., & Klawonn, F. (2024). A Novel Tool for the Rapid and Transparent Verification of Reference Intervals in Clinical Laboratories. Journal of clinical medicine, 13(15), 4397. <https://doi.org/10.3390/jcm13154397>.
