---
title: "Reference Limits with Limit of Detection (LOD)"
author: "Sandra Klawitter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reference Limits with Limit of Detection (LOD)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4.5
)
```

## Introduction
Laboratory measurements frequently encounter values below lower limit of detection or truncated distributions. 

## Usage
### 1. Estimating Reference Limits with LOD
The reflimLOD.MLE() function adjusts reference limit calculations when a portion of the dataset is below a specified detection threshold:
```{r}
library(reflimR.expand)

set.seed(42)
# Generate synthetic biomarker data
raw_data <- rnorm(800, mean = 15, sd = 4)

# Define a Limit of Detection (LOD)
lod_value <- 10

# Separate values above LOD and count values below LOD
measured_values <- raw_data[raw_data >= lod_value]
n_lod <- sum(raw_data < lod_value)

# Calculate reference limits using MLE
res_lod <- reflimLOD.MLE(
  measured.values = measured_values,
  lod = lod_value,
  n.lod = n_lod,
  lambda = 0
)

# Inspect output structure
res_lod
```
### 2. Visualization

You can visualize the distribution and estimated reference bounds directly:
```{r}
# Plotting a Basic Histogram
hist(res_lod$selected.values,
     main = "Reference Interval with LOD Data",
     xlab = "Biomarker Level",
     col = "lightblue",
     border = "white")

# Use vectorized functions to compute and superimpose density curves
f_dens <- Vectorize(function(x) {
  area <- (length(res_lod$selected.values) + res_lod$n.lod) * 1  # 自动计算步长面积
  area * dens.box.cox.inv(x, mu = res_lod$mu.log, sigma = res_lod$sigma.log, lambda = res_lod$lambda)
})

curve(f_dens, add = TRUE, col = "red", lwd = 2)
```

## References

Klawonn, F., Hoffmann, G., Holdenrieder, S., & Trulson, I. (2024). reflimLOD: A Modified reflimR Approach for Estimating Reference Limits with Tolerance for Values Below the Lower Limit of Detection (LOD). Stats, 7(4), 1296-1314. https://doi.org/10.3390/stats7040075 
