---
title: "Indirect Reference Interval Estimation using Decision Trees (rpart)"
author: "Sandra Klawitter, Weihao Qiu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Indirect Reference Interval Estimation using Decision Trees (rpart)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
This vignette demonstrates regression tree partitioning using rpart for indirect reference interval estimation across continuous and categorical covariates (e.g., age and sex).

## Usage
Below is a standard workflow demonstration using the generate_rpart() function for regression tree partitioning and visualization.

```{r setup, message = FALSE}
library(reflimR.expand)
library(reflimR)
library(rpart)
```

We can estimate age- and sex-dependent reference intervals using partitioning decision trees on clinical laboratory data:

```{r example}
# Load example dataset
data("livertests", package = "reflimR")

# Fit decision tree model to partition ALT by Age and Sex
fit_tree <- generate_rpart(
  df = livertests,
  analyte = "ALT",
  tree_cp = 0.01,
  tree_minsplit = 30
)

# Print tree summary
print(fit_tree)
```

## References

Klawitter, Sandra, Böhm, Johannes, Tolios, Alexander and Gebauer, Julian E.. "Automated sex and age partitioning for the estimation of reference intervals using a regression tree model" Journal of Laboratory Medicine, vol. 48, no. 5, 2024, pp. 223-237. https://doi.org/10.1515/labmed-2024-0083
