---
title: "Indirect Reference Interval Estimation using Sliding Window and Decision Trees (rpart)"
author: "Sandra Klawitter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Indirect Reference Interval Estimation using Sliding Window and Decision Trees (rpart)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
The rpart decision tree partitioning for indirect reference interval estimation across continuous covariates (e.g., age).

## Usage
Below is a standard workflow demonstration for using sliding window decision tree algorithms.

```{r setup, message = FALSE}
library(reflimR.expand)
library(reflimR)
```

We can estimate age- and sex-dependent reference intervals using partitioning decision trees:

```{r example, eval = FALSE}
# Load example dataset from reflimR
data("livertests", package = "reflimR")

# Run decision tree sliding window algorithm (assuming age and analyte columns)
# res <- rpart_window(livertests, analyte = "ALT", covariate = "age")

# Print summary
# summary(res)
```

## References

Klawitter, Sandra, Böhm, Johannes, Tolios, Alexander and Gebauer, Julian E.. "Automated sex and age partitioning for the estimation of reference intervals using a regression tree model" Journal of Laboratory Medicine, vol. 48, no. 5, 2024, pp. 223-237. https://doi.org/10.1515/labmed-2024-0083
