---
title: "Sliding Window Methods for Age-Dependent Reference Intervals"
author: "Sandra Klawitter, Weihao Qiu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sliding Window Methods for Age-Dependent Reference Intervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4.5
)
```

## Introduction

Biological parameters often change continuously across covariates such as age. Instead of creating arbitrary discrete age groups, sliding.reflim() applies a moving window to compute continuous local reference limits.

## Usage

### 1. Standard Sliding Window (sliding.reflim)
```{r}
library(reflimR)
library(reflimR.expand)

set.seed(123)
# Generate synthetic age data (18 to 80 years)
age <- runif(1000, min = 18, max = 80)

# Generate biomarker values with an age-dependent upward trend
biomarker <- 40 + 0.12 * age + rnorm(1000, mean = 0, sd = 2.5)

# Calculate sliding reference limits
res_sliding <- sliding.reflim(
  x = biomarker,
  covariate = age,
  window.size = 20,
  step.width = 2,
  n.min = 39
)


# Inspect output structure
head(res_sliding)
```

The companion function draw.sliding.reflim() visualizes local reference limits. You can plot results aligned by local median covariate (use.mean = FALSE) or mean covariate (use.mean = TRUE):

```{r}
draw.sliding.reflim(res_sliding, use.mean = FALSE)
```

### 2. Weighted Sliding Window (sliding.w.reflim)
For datasets with sparse regions or non-uniform age distributions, localized kernel weights can be applied via weighted sliding functions (sliding.w.reflim) to assign higher weights to observations near the center of the window:

```{r weighted_sliding, warning = FALSE, message = FALSE}
# Run weighted sliding window and display the plot directly
sliding.w.reflim(
  x = biomarker,
  t = age,
  distribution = "gaussian",
  standard_deviation = 5,
  n.min = 39,
  verbose = FALSE
)
```
