---
title: "Indirect Reference Interval Estimation using Sliding Window and Decision Trees (rpart)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Indirect Reference Interval Estimation using Sliding Window and Decision Trees (rpart)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
The rpart sliding window methods integrate decision tree partitioning (rpart) with sliding window algorithms for indirect reference interval estimation across continuous covariates (e.g., age). Ported from the AdRI and AdRI_rpart repositories, these functions enable continuous, non-linear partitioning of reference limits.

## Basic Usage
Below is a standard workflow demonstration for using sliding window decision tree algorithms.

```{r setup, message = FALSE}
library(reflimR.expand)
library(reflimR)
```

### Example Usage
We can estimate age-dependent reference intervals using partitioning decision trees:

```{r example, eval = FALSE}
# Load example dataset from reflimR
data("livertests", package = "reflimR")

# Run decision tree sliding window algorithm (assuming age and analyte columns)
# res <- rpart_window(livertests, analyte = "ALT", covariate = "age")

# Print summary
# summary(res)
```

## References

- Sliding window algorithms original sources:
  - [AdRI (window.R)](https://github.com/SandraKla/AdRI/blob/master/R/window.R)
  - [AdRI_rpart (rpart.R)](https://github.com/SandraKla/AdRI_rpart/blob/main/rpart.R)