---
title: "Standardization of Laboratory Data using zlog"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standardization of Laboratory Data using zlog}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
The zlog() function and its related utilities provide methods for standardizing and normalizing laboratory measurement data using reference limits via $z$-score and logarithmic transformations. Ported from the VeRIf repository, zlog helps make laboratory values comparable across different analytes or instruments.

## Basic Usage
Below is a basic demonstration of applying zlog transformations.

```{r setup, message = FALSE}
library(reflimR.expand)
library(reflimR)
```

### Example Usage
We can transform laboratory measurements based on defined lower and upper reference limits:

```{r example, eval = FALSE}
# Example analyte measurements
vals <- c(10, 15, 20, 25, 30, 100)

# Define reference limits (e.g., Lower Limit = 12, Upper Limit = 25)
limits <- c(lower = 12, upper = 25)

# Calculate zlog values
z_scores <- zlog(vals, limits = limits)

# Print standardized zlog scores
print(z_scores)
```

## References

- Algorithm original source: [VeRIf (functions.R)](https://github.com/SandraKla/VeRIf/blob/main/functions.R)